---
title: "Assignment 1: Writeup"
author: "Kyle Kurkela"
date: "10/11/2020"
output: pdf_document
---

```{r, echo=FALSE, warning=FALSE}
shh <- suppressPackageStartupMessages
shh(library(tidyverse))
shh(library(kableExtra))
shh(library(nlme))

vars_of_interest <- c('MATHINT', 'MATHUTIL', 'MATHCONF', 'COGLEARN', 'METACOG')

# descriptives
school_counts <- read_rds('../Assignment1/school_counts.rds')
df <- read_rds('../Assignment1/data.rds')
nstudents <- nrow(df)
cluster.stats <- read_rds('../Assignment1/clusters_stats.rds')

# models
Model1 <- read_rds('Model1.Rds')
Model2a <- read_rds('Model2a.Rds')
Model3 <- read_rds('Model3.Rds')
```

## Descriptive Statistics

The PISA data had observations from `r 4997` Russian students from `r nrow(school_counts)` schools. Each school contributed, on average, `r round(cluster.stats$mean, 2)` students with a standard deviation of `r round(cluster.stats$sd, 2)`. The number of students per school ranged from 1 to 35, with the half of schools contributing more than `r cluster.stats$median` students. See Figure 1 for a histogram of the distribution the number of students per school. Descriptive statistics for the following variables of interest are summarised in Table 1: Math interest, Math utility, Math confidence, Cognitive learning strategy, and Metacognitive learning strategy.

## Modeling

Three different models were run to investigate the effect of Math interest, Math utility, and Math confidence on Metacognitive learning strategy: Model 1: a model a random intercept nested within schools and fixed effects, Model 2: a model with a random intercept and random slopes for Math utility and Math confidence nested within schools, and Model 3: a model with random intercept and random slopes for Math utility and math confidence nested within schools using a variance component covariance structure (see Table 2). 

Models 1, 2, and 3 all revelaed similar effects of ...

## Figures

```{r echo=FALSE, fig.align='center', fig.height=3, fig.width=5}
Fig1 <- read_rds('../Assignment1/Fig1.rds')
Fig1
```
  
```{r calculate_fixedEff, echo=FALSE, message=FALSE, warning=FALSE}
# Model 1

# compute confidence intervals around fixed effect estimates
Model1 %>%
confint(method = 'profile') %>%
  as_tibble(rownames = 'parameter') %>%
  filter(str_detect(parameter, 'MATH')) -> FixedEf.ConfInt

# fixed effect estimates
Model1 %>%
  fixed.effects() %>%
  as_tibble(rownames = 'parameter') %>%
  filter(str_detect(parameter, 'MATH')) -> FixedEf.Est

left_join(FixedEf.Est, FixedEf.ConfInt, by = 'parameter') -> Model1.FixedEffects

# Model 2

# compute confidence intervals around fixed effect estimates
Model2a %>%
confint(method = 'profile') %>%
  as_tibble(rownames = 'parameter') %>%
  filter(str_detect(parameter, 'MATH')) -> FixedEf.ConfInt

# fixed effect estimates
Model2a %>%
  fixed.effects() %>%
  as_tibble(rownames = 'parameter') %>%
  filter(str_detect(parameter, 'MATH')) -> FixedEf.Est

left_join(FixedEf.Est, FixedEf.ConfInt, by = 'parameter') -> Model2.FixedEffects

# Model 3

# compute confidence intervals around fixed effect estimates
# apparently you cannot use the "stats::confint" function on lme objects.
# the package authors suggest using "nlme::intervals"
Model3 %>%
  intervals(which = 'fixed') -> Conf.Intervals

Conf.Intervals$fixed %>% 
  as_tibble(rownames = 'parameter') %>%
  filter(str_detect(parameter, 'MATH')) %>%
  rename(`2.5 %` = lower,
         `97.5 %` = upper,
         value = est.) -> Model3.FixedEffects

bind_rows(Model1.FixedEffects, Model2.FixedEffects, Model3.FixedEffects, .id = 'Model') -> AllModels.df
```
  
## Tables

```{r echo=FALSE}
Tbl1 <- read_rds('../Assignment1/Tbl1.rds')

options(knitr.kable.NA = '')

Tbl1 %>%
  kable(format = 'latex',
        col.names = c('','', vars_of_interest[1:length(vars_of_interest)-1], ''), 
        caption = 'Mean, standard deviation, and correlation of Level 1 variables',
        digits = 2, row.names = F,
        align = c('l','c','c','c','c','c','c'),
        booktabs = T) %>%
  kable_styling(latex_options = 'HOLD_position',
                position = 'center') %>%
  add_header_above(header = c('Variables', 'Mean (SD)',
                              'Correlation' = 4, 'ICC')) %>%
  footnote(general = 'N = 4997 students. SD = standard deviation. ICC = intraclass correlation. Means, standard deviations, and correlations were computed ignoring the clusted structure. MATHINT = Math interest, MATHUTIL = Math utility, MATHCONF = Math confidence, COGLEARN = Cognitive learning strategy, METACOG = Metacognitive learning strategy.', threeparttable = T, footnote_as_chunk = T)
```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
summary(Model1) -> Model1.Summary

Model1.Summary$coefficients %>%
  as_tibble(rownames = 'parameter') %>%
  mutate(Estimate = signif(Estimate, digits = 3),
         `Std. Error` = signif(`Std. Error`, digits = 2),
         df = signif(df, digits = 4),
         `t value` = signif(`t value`, 4)) %>%
  transmute(parameter = parameter,
            Estimate_se = str_glue('{Estimate} ({`Std. Error`})'),
            t_df = str_glue('{`t value`} ({df})'),
            `p value` = 'p < .001') -> MixedModelTbl.Pt1

confint(Model1) -> ConfidenceIntervals

ConfidenceIntervals %>%
  as_tibble(rownames = 'parameter') %>%
  filter(str_detect(parameter, 'sig', negate = T)) -> MixedModelTbl.Pt2

left_join(MixedModelTbl.Pt1, MixedModelTbl.Pt2, by = 'parameter') %>%
  mutate(across(ends_with('%'), signif, digits = 3)) -> MixedModelTbl

# Random Effects
VarIntercept <- merTools::REsdExtract(Model1)^2
ResidualVar <- Model1.Summary$sigma^2

tibble(parameter = c('.sig01', '.sigma'), estimate = c(VarIntercept, ResidualVar)) -> TmpTbl

ConfidenceIntervals %>%
  as_tibble(rownames = 'parameter') %>%
  filter(str_detect(parameter, 'sig')) -> TmpTbl2
  
left_join(TmpTbl, TmpTbl2) %>%
  mutate(across(where(is.double), signif, digits = 3)) -> SigmaTbl

MixedModelTbl %>%
  bind_rows(SigmaTbl) %>%
  transmute(parameter = parameter,
            Estimate_se = Estimate_se,
            t_df = t_df,
            `p value` = `p value`,
            `95 % confidence interval` = str_glue('[{`2.5 %`}, {`97.5 %`}]')) -> MixedModelTbl
  kable(format = 'latex',
        col.names = c('Fixed effects', 'Estimate (SE)', 't (df)', 'p value', '95 % confidence interval'), 
        caption = 'Estimated mixed model -- Model 1', 
        row.names = F,
        booktabs = T) %>%
  kable_styling(latex_options = 'HOLD_position',
                position = 'center') %>%
  footnote(general = '', threeparttable = T, footnote_as_chunk = T)
```

