---
title: "Assignment 2: Writeup"
author: "Kyle Kurkela"
date: "11/12/2020"
output: word_document
---

```{r, echo=FALSE, warning=FALSE}
shh <- suppressPackageStartupMessages
shh(library(tidyverse))
shh(library(kableExtra))
shh(library(nlme))

vars_of_interest <- c('MATHINT', 'MATHUTIL', 'MATHCONF', 'COGLEARN', 'METACOG')

# descriptives
school_counts <- read_rds('../Assignment1/school_counts.rds')
df <- read_rds('../Assignment1/data.rds')
nstudents <- nrow(df)
cluster.stats <- read_rds('../Assignment1/clusters_stats.rds')

# models
Model1 <- read_rds('Model1.Rds')
Model2a <- read_rds('Model2a.Rds')
Model3 <- read_rds('Model3.Rds')
```

## Descriptive Statistics

The PISA data had observations from `r 4997` Russian students from `r nrow(school_counts)` schools. Each school contributed, on average, `r round(cluster.stats$mean, 2)` students with a standard deviation of `r round(cluster.stats$sd, 2)`. The number of students per school ranged from 1 to 35, with the half of schools contributing more than `r cluster.stats$median` students. See Figure 1 for a histogram of the distribution the number of students per school. Descriptive statistics for the following variables of interest are summarised in Table 1: Math interest, Math utility, Math confidence, Cognitive learning strategy, and Metacognitive learning strategy.

## Modeling

Three different models were run to investigate the effect of Math interest, Math utility, and Math confidence on Metacognitive learning strategy: Model 1: a model with a random intercept and fixed effects, Model 2: a model with a random intercept and random slopes for Math utility and Math confidence, and Model 3: a model with random intercept and random slopes for Math utility and math confidence nested within schools using a variance component covariance structure (see Table 2). 

Models 1, 2, and 3 all revelaed similar effects of ...

## Figures

```{r echo=FALSE, fig.align='center', fig.height=3, fig.width=5}
Fig1 <- read_rds('../Assignment1/Fig1.rds')
Fig1
```
  
## Tables

```{r echo=FALSE}
Tbl1 <- read_rds('../Assignment1/Tbl1.rds')

options(knitr.kable.NA = '')

# Tbl1 %>%
#   kable(col.names = c('','', vars_of_interest[1:length(vars_of_interest)-1], ''), 
#         caption = 'Mean, standard deviation, and correlation of Level 1 variables',
#         digits = 2, row.names = F,
#         align = c('l','c','c','c','c','c','c'),
#         booktabs = T) %>%
#   kable_styling(latex_options = 'HOLD_position',
#                 position = 'center') %>%
#   add_header_above(header = c('Variables', 'Mean (SD)',
#                               'Correlation' = 4, 'ICC')) %>%
#   footnote(general = 'N = 4997 students. SD = standard deviation. ICC = intraclass correlation. Means, standard deviations, and correlations were computed ignoring the clusted structure. MATHINT = Math interest, MATHUTIL = Math utility, MATHCONF = Math confidence, COGLEARN = Cognitive learning strategy, METACOG = Metacognitive learning strategy.', threeparttable = T, footnote_as_chunk = T)
# ```

