---
title: "Assignment 2: Code"
author: "Kyle Kurkela"
date: "11/06/2020"
output: html_document
---

Below is the code used to run the analyses for Assignment 2.

Like for assignment 1, I choose to analyze the `PISA: Russia` dataset.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(lme4)
library(performance)
library(lmerTest)
library(nlme)

df <- read_csv('../data/pisaRUS.csv', col_types = cols())

# select at least 5 continuous level 1 variables
df %>%
  select(SCHOOLID, MATHINT, MATHUTIL, MATHCONF, COGLEARN, METACOG) -> df
```

### Select a dependent variable. Choose at least 2 level-1 predictors.

For my analysis, I will use metacognitive learning strategy (`METCOG`) as my dependent variable and Math interest (`MATHINT`), Math utility (`MATHUTIL`), and Math confidence (`MATHCONF`) as my predictors.

```{r}
# select the dv and the predictors
df %>%
  select(SCHOOLID, METACOG, MATHINT, MATHUTIL, MATHCONF) -> df
```

### Model 1: Random Intercepts and Fixed Slopes

The first model will contain random intercepts by school and fixed slopes. In equation form:

**Multilevel Form**:
  
Level 1:  
$METACOG_{i,j} = \beta_{0j} + \beta_{1,j} MATHINT + \beta_{2,j} MATHUTIL + \beta_{3,j} MATHCONF$  
$e_{ij} \thicksim N(0,\sigma^2)$  
  
Level 2:  
$\beta_{0j} = \gamma_{00} + u_{0j}$  
$\beta_{1j} = \gamma_{10}$  
$\beta_{2j} = \gamma_{20}$  
$\beta_{3j} = \gamma_{30}$  
  
**Mixed Model Form**:  
  
$$METACOG_{ij} = \gamma_{00} + \gamma_{10} MATHINT_{ij} + \gamma_{20} MATHUTIL_{i,j} + \gamma_{30} MATHCONF_{ij} + u_{0j} + e_{ij}$$

**Parameters**:
Fixed effects coefficients:  
- $\gamma_{00}$  
- $\gamma_{10}$  
- $\gamma_{20}$  
- $\gamma_{30}$  
Variance of the random intercept:  
- $Var(u_{0j}) = \tau_{00}$  
Variance of the level-1 residual:  
- $Var(e_{ij} = \sigma^2)$  

6 Total Parameters

```{r}
# Run Model 1, making sure to use lmerTest version of `lmer` function
Model1 <- lmerTest::lmer(METACOG ~ MATHINT + MATHUTIL + MATHCONF + (1|SCHOOLID), data = df)

write_rds(Model1, 'Model1.Rds')

summary(Model1)
```

### Model 2: Random Intercepts and Random Slopes

The second model will contain random intercepts by school and random slopes for MATHINT, MATHUTIL, and MATHCONF. In equation form:

**Multilevel Form**:
  
Level 1:  
$METACOG_{i,j} = \beta_{0j} + \beta_{1,j} MATHINT + \beta_{2,j} MATHUTIL + \beta_{3,j} MATHCONF$  
$e_{ij} \thicksim N(0,\sigma^2)$  
  
Level 2:  
$\beta_{0j} = \gamma_{00} + u_{0j}$  
$\beta_{1j} = \gamma_{10} + u_{1j}$  
$\beta_{2j} = \gamma_{20} + u_{2j}$  
$\beta_{3j} = \gamma_{30} + u_{3j}$  
  
$\boldsymbol{u}_{j} \thicksim MVN(\boldsymbol{0}, \boldsymbol{T})$
  
**Mixed Model Form**:  
  
$$METACOG_{ij} = \gamma_{00} + \gamma_{10} MATHINT_{ij} + \gamma_{20} MATHUTIL_{i,j} + \gamma_{30} MATHCONF_{ij} + u_{0j} + u_{0j} + u_{1j} MATHINT_{ij} + u_{2j} MATHUTIL_{ij} + u_{3j} MATHCONF_{ij}  + e_{ij}$$

**Parameters**:
Fixed effects coefficients:  
- $\gamma_{00}$  
- $\gamma_{10}$  
- $\gamma_{20}$  
- $\gamma_{30}$  
Variance of the level 2 random effects:
- $\boldsymbol{T}$ = (4 x 4) covariance matrix with 10 paramaters  
Variance of the level-1 residual:  
- $\sigma^2$  

15 Total Parameters

```{r}
# Run Model 1, making sure to use lmerTest version of `lmer` function
Model2 <- lmerTest::lmer(METACOG ~ MATHINT + MATHUTIL + MATHCONF + (MATHINT + MATHUTIL + MATHCONF| SCHOOLID), data = df)

summary(Model2)
```

The model failed to converge! Lets make the model simpler by dropping one of the random effects. The MATHINT random slopes seems to have the smallest variance (0.00296), so lets try dropping that one:

```{r}
Model2a <- lmerTest::lmer(METACOG ~ MATHINT + MATHUTIL + MATHCONF + (MATHUTIL + MATHCONF | SCHOOLID), data = df)

write_rds(Model2a, 'Model2a.Rds')

summary(Model2a)
```

Success! The model has converged.  

### Model 3: Random Intercept and Random slopes with Variance Components (VC) covariance structure

The second model will contain random intercepts by school and random slopes for MATHUTIL and MATHCONF (see Model 2a). In equation form:

**Multilevel Form**:
  
Level 1:  
$METACOG_{i,j} = \beta_{0j} + \beta_{1,j} MATHINT + \beta_{2,j} MATHUTIL + \beta_{3,j} MATHCONF$  
$e_{ij} \thicksim N(0,\sigma^2)$  
  
Level 2:  
$\beta_{0j} = \gamma_{00} + u_{0j}$  
$\beta_{1j} = \gamma_{10}$  
$\beta_{2j} = \gamma_{20} + u_{2j}$  
$\beta_{3j} = \gamma_{30} + u_{3j}$  
  
Variance components (VC) covariance structure:  
  
$$\begin{bmatrix} u_0j \\ u_{2j} \\ u_{3j} \end{bmatrix} \thicksim MVN(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, \begin{bmatrix} \tau_{00} & & \\ 0 & \tau_{11} & \\ 0 & 0 & \tau_{33} \end{bmatrix})$$
  
**Mixed Model Form**:  
  
$$METACOG_{ij} = \gamma_{00} + \gamma_{10} MATHINT_{ij} + \gamma_{20} MATHUTIL_{i,j} + \gamma_{30} MATHCONF_{ij} + u_{0j} + u_{0j} + u_{2j} MATHUTIL_{ij} + u_{3j} MATHCONF_{ij}  + e_{ij}$$

**Parameters**:
Fixed effects coefficients:  
- $\gamma_{00}$  
- $\gamma_{10}$  
- $\gamma_{20}$  
- $\gamma_{30}$  
Variance of the level 2 random effects:
- $\boldsymbol{T}$ = (3 x 3) covariance matrix with 3 paramaters  
Variance of the level-1 residual:  
- $\sigma^2$  

8 Total Parameters

```{r paged.print=FALSE}
library(nlme)
Model3 <- lme(METACOG ~ MATHINT + MATHUTIL + MATHCONF, 
              random = list(SCHOOLID = pdDiag(~1 + MATHUTIL + MATHCONF)), data = df)

write_rds(Model3, 'Model3.Rds')

summary(Model3)
```
```{r}
# Get the Variance-Covariance matrix
getVarCov(Model3)
```

Looks like it worked!!
