---
title: "Assignment 1: Code"
author: "Kyle Kurkela"
date: "10/11/2020"
output: html_document
---

Below is the code used to run the descriptive analysis for Assignment 1.

I choose to analyze the `PISA: Russia` dataset.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(lme4)
library(performance)

df <- read_csv('pisaRUS.csv', col_types = cols())

# select at least 5 continuous level 1 variables
df %>%
  select(SCHOOLID, MATHINT, MATHUTIL, MATHCONF, COGLEARN, METACOG) -> df
```

## Cluster Descriptives

Code used to get descriptive statistics for the number of students per cluster (i.e., School).

```{r}
nrow(df)
```

Each row of the `df` corresponds to a student. There are a total of `r nrow(df)` students in the `PISA: Russia` dataset.  

```{r descriptives}
df %>%
  count(SCHOOLID) -> school_counts

descriptives <- c('n', 'min', 'max', 'mean', 'median', 'sd')

psych::describe(x = school_counts) %>%
  filter(vars == 2) %>%
  select(all_of(descriptives)) %>%
  magrittr::set_rownames('frequency') -> cluster.stats

cluster.stats
```

```{r}
ggplot(school_counts, aes(n)) +
  geom_histogram(bins = 30) +
  labs(y = 'Frequnecy', x = 'Number of Students',
       title = 'Figure 1: Number of Students in Each School',
       subtitle = 'PISA: Russia') +
  theme_classic() -> Fig1

Fig1
``` 

## Level 1 Descriptives

Code used to generate a table of descriptive statistics for the Level 1 variables.  

```{r}
vars_of_interest <- c('MATHINT', 'MATHUTIL', 'MATHCONF', 'COGLEARN', 'METACOG')
nvars <- length(vars_of_interest)

options(knitr.kable.NA = '')

# Level 1: Mean and Standard Deviation calculation
df %>%
  select(-SCHOOLID) %>%
  psych::describe() %>%
  select(mean, sd) %>%
  as_tibble(rownames = 'vars') %>%
  mutate(across(where(is.double), .fns = round, digits = 2)) %>%
  transmute(vars = vars,
            mean_sd = str_glue('{mean} ({sd})')) -> lvl1.summary

# Level 1: Correlation Matrix Calculation
corrr::correlate(df %>% select(all_of(vars_of_interest)), quiet = TRUE) %>%
  corrr::shave() %>%
  select(-where(~is.na(.x) %>% all())) -> lvl1.corrs

# Level 1: ICC calculations
FIT.list <- vector(mode = 'list', length = nvars)
ICC.list <- vector(mode = 'double', length = nvars)
for(v in 1:nvars){
  FORM <- str_glue('{vars_of_interest[v]} ~ 1 + (1|SCHOOLID)')
  FIT <- lmer(as.formula(FORM), data = df)
  FIT.list[[v]] <- FIT
  ICC.list[v] <- icc(FIT) %>% as.double() %>% unique()
}

# Joining into a table
left_join(lvl1.summary, lvl1.corrs, by = c('vars' = 'rowname')) %>%
  add_column(ICC = ICC.list) -> Tbl1
```

```{r}
# MATHINT
summary(FIT.list[[1]])
```

```{r}
# MATHUTIL
summary(FIT.list[[2]])
```

```{r}
# METHCONF
summary(FIT.list[[3]])
```

```{r}
# COGLEARN
summary(FIT.list[[4]])
```

```{r}
# METACOG
summary(FIT.list[[5]])
```

```{r}
# Pretty Printing of Table
Tbl1 %>%
  kable(col.names = c('','', vars_of_interest[1:length(vars_of_interest)-1], ''), 
        caption = 'Table 1: Mean, standard deviation, and correlation of Level 1 variables',
        digits = 2, row.names = F,
        align = c('l','c','c','c','c','c','c')) %>%
  kable_styling() %>%
  add_header_above(header = c('Variables', 'Mean (SD)', 'Correlation' = 4, 'ICC')) -> Tbl1.pretty

Tbl1.pretty
```

