---
title: "Assignment 1: Writeup"
author: "Kyle Kurkela"
date: "10/11/2020"
output: pdf_document
---

```{r, echo=FALSE, warning=FALSE}
shh <- suppressPackageStartupMessages
shh(library(tidyverse))
shh(library(kableExtra))

vars_of_interest <- c('MATHINT', 'MATHUTIL', 'MATHCONF', 'COGLEARN', 'METACOG')

school_counts <- read_rds('school_counts.rds')
df <- read_rds('data.rds')
nstudents <- nrow(df)
cluster.stats <- read_rds('clusters_stats.rds')
```

## Descriptive Statistics

The PISA data had observations from `r 4997` Russian students from `r nrow(school_counts)` schools. Each school contributed, on average, `r round(cluster.stats$mean, 2)` students with a standard deviation of `r round(cluster.stats$sd, 2)`. The number of students per school ranged from 1 to 35, with the half of schools contributing more than `r cluster.stats$median` students. See Figure 1 for a histogram of the distribution the number of students per school. Descriptive statistics for the following variables of interest are summarised in Table 1: Math interest, Math utility, Math confidence, Cognitive learning strategy, and Metacognitive learning strategy.

## Figures

```{r echo=FALSE, fig.align='center', fig.height=3, fig.width=5}
Fig1 <- read_rds('Fig1.rds')
Fig1
```
  
  
## Tables

```{r echo=FALSE}
Tbl1 <- read_rds('Tbl1.rds')

options(knitr.kable.NA = '')

Tbl1 %>%
  kable(format = 'latex',
        col.names = c('','', vars_of_interest[1:length(vars_of_interest)-1], ''), 
        caption = 'Table 1: Mean, standard deviation, and correlation of Level 1 variables',
        digits = 2, row.names = F,
        align = c('l','c','c','c','c','c','c'),
        booktabs = T) %>%
  kable_styling(latex_options = 'HOLD_position',
                position = 'center') %>%
  add_header_above(header = c('Variables', 'Mean (SD)',
                              'Correlation' = 4, 'ICC')) %>%
  footnote(general = 'N = 4997 students. SD = standard deviation. ICC = intraclass correlation. Means, standard deviations, and correlations were computed ignoring the clusted structure. MATHINT = Math interest, MATHUTIL = Math utility, MATHCONF = Math confidence, COGLEARN = Cognitive learning strategy, METACOG = Metacognitive learning strategy.', threeparttable = T, footnote_as_chunk = T)
```

